---
title: "Predicting the Prices of Homes in Boston"
author: "Multiple Linear Regression INSOFE Lab Session"
date: "18 June 2017"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
---

**NOTE** Before starting this assignment please remember to clear your environment, you can do that by running the following code chunk

```{r}

rm(list = ls(all=TRUE))

```

# Agenda 

* Get the data

* Ask an interesting question

* Explore the data

* Data Pre-processing

* Model the data

* Evaluation


# Reading & Understanding the Data

* Make sure the dataset is located in your current working directory

```{r}



```

* Use the str() function to get a feel for the dataset.

```{r}



```

* The dataset has 506 rows and 14 columns.

* The column/variable names' explanation is given below:

1) __CRIM :__ Per capita Crime rate by town

2) __ZN :__ Proportion of residential land zoned for lots over 25,000 sq.ft.

3) __INDUS :__ Proportion of non-retail business acres per town

4) __CHAS :___ Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)

5) __NOX :__ nitric oxides concentration (parts per 10 million)

6) __RM :__ average number of rooms per dwelling

7) __AGE :__ proportion of owner-occupied units built prior to 1940

8) __DIS :__ weighted distances to five Boston employment centres

9) __RAD :__ index of accessibility to radial highways

10) __TAX :__ full-value property-tax rate per $10,000

11) __PTRATIO :__ pupil-teacher ratio by town

12) __B :__ 1000(Bk - 0.63)^2 where Bk is the proportion of African-Americans by town

13) __LSTAT :__ Percentage of the population in the lower economic status 

14) __MEDV  :__ Median value of owner-occupied homes in multiples of $1000


* Take a look at the data using the "head()" and "tail()" functions

```{r}




```


# Ask and Explore interesting Questions

1) __How does the proportion of African-Americans by town relate to the Median Value of owner-occupied homes?__

2) __How does the nitric oxides concentration in ppm in the air relate to the Median Value of owner-occupied homes?__

3) ????

4) ????

* So, let's get ready to predict the prices of houses!!

# Data Pre-Processing - I

## Missing Values imputation

* Find out the number of missing values in the dataset

* Impute the missing values using the "centralImputation()" function from the DMwR package

```{r}




```

# Exploratory Analysis

## Summary Statistics

* Understand the distribution of various variables in the datset using the "summary()" function

```{r}



```

## Scatter Plots

* A few bi-variate relationships are plotted below, but you are encouraged to explore the dataset in more detail

```{r fig.height= 8, fig.width = 9}




```



## Correlation Plot

* Let's have a look at the various correlations between the variables in the dataset

```{r fig.height= 8, fig.width = 9}




```


# Data Pre-processing - II

* Today we will standardize the data __after__ splitting the data into train and test sets

## Train/Test Split

* 70/30 - Train/Test split

```{r}



```


## Standardizing the Data

* We will use the Caret pakcage to standardize the data after the split using the __"preProcess()"__ function

* It saves the metrics such as mean and standard deviation used for calculating the standardized value by creating a model object

* We can then use the model object in the __"predict()"__ function to standardize any other unseen dataset with the same distribuiton and variables

```{r}

library(caret)

# The "preProcess()" function creates a model object required for standardizing unseen data

# Do not standardize the target variable



# The predict() function is used to standardize any other unseen data



```


# Modelling the Data

## Basic Model

* The "." adds all the variables other than the response variable while building the model.

```{r}





```

## stepAIC model

* "stepAIC()" is a function in the MASS package

* stepAIC uses AIC (Akaike information criterion) to either drop variables ("backward" direction) or add variables ("forward" direction) from the model

```{r}


# Here we use the "backward" direction as model_basic used all the variables to build the regression model


```

## Modifying the Model with the VIF

**Variance Inflation Factor :**

$$VIF_{k} = \dfrac{1}{1 - R_{k}^2}$$

$R_{k}^2$ is the R^2-value obtained by regressing the kth predictor on the remaining predictors. VIF gives us an idea of multi-collinearity

* Every explanatory variable would have a VIF score

* A VIF > 4 means that there are signs of multi-collinearity and anything greater than 10 means that an explanatory variable should be dropped

* We use the "vif()" function from the car package. 

```{r}



```

* After applying the stepAIC, the VIF values have slightly reduced, but the variables "RAD" and "TAX" have VIF values higher than 4

* Let's take a look at the correlation between the "RAD" and "TAX" variables

```{r}



```

* The correlation coefficient is extremely high between the "RAD" and "TAX" variables

* let's now remove the "TAX" variable, as it is the lesser significant of the two

* Build another model without the "TAX" variable, and take a look at the VIFs


```{r}



```


# Evaluation and Selection of Model

## Picking the Right Model

* The third model built after verifying the vif scores has a similar adjusted R^2 score compared to the previous models with significantly lower no. of explanatory variables and inter-variable interactions.

* The VIF values of the predictors in the third model are lower when compared to the the other two models

* Due to the above two reasons we pick the third model

# Communication

## Prediction

Predict the Housing prices of the unseen boston housing data, using the chosen model.

```{r}


```

## Performance Metrics

Once we choose the model we have to report performance metrics on the test data. We are going to report three error metrics for regression.

### Error Metrics for Regression

* Mean Absolute Error (MAE)

Create a function called mae that measures the mean absolute error, given the actual and predicted points.

$$MAE = \dfrac{1}{n}\times|\sum_{i = 1}^{n}y_{i} - \hat{y_{i}}|$$

```{r}



```

* Mean Squared Error (MSE)

Create a function called mse that measures the mean squared error, given the actual and predicted points.

$$MSE = \dfrac{1}{n}\times(\sum_{i = 1}^{n}y_{i} - \hat{y_{i}})^2$$

```{r}



```

* Root Mean Squared Error (RMSE)

Create a function called rmse that measures the root mean squared error, given the actual and predicted points.

$$RMSE = \sqrt{\dfrac{1}{n}\times(\sum_{i = 1}^{n}y_{i} - \hat{y_{i}})^2}$$

```{r}



```


* Mean Absolute Percentage Error (MAPE)

Create a function called mae that measures the mean absolute error, given the actual and predicted points.

$$MAPE = \dfrac{100}{n}\times\mid\dfrac{\sum_{i = 1}^{n}y_{i} - \hat{y_{i}}}{y_{i}}\mid$$

```{r}




```


### Report Performance Metrics

* Report performance metrics obtained by using the chosen model on the test data

```{r}




```



















































