---
title: "Multiple Linear Regression on Customer Data"
author : "Insofe Lab Session"
date: "June 11 , 2017"
output:
  html_document:
    toc_depth: 3
    toc_float: true
  

---
```{r}
# Go through the data and understand the attributes to get the info of the variables.
```

Write a code to clear the environment if needed
```{r}
rm(list = ls(all=TRUE))
```

Write a code to set the working directory.
```{r }
setwd("C:/Users/kiranpatro/Desktop/INSOFE notes/assignments lab/20170618-batch29-cse7302c-multiple-linear-regression-kiran-master")
```
Think and load the libraries as and when required in this place only as best practice
```{r}

```
Write a code to read the csv file as "data" as per required
```{r eval=TRUE }
customer_data=read.csv("customer_data.csv")

```
Write a code to know the names of the attributes
```{r}
names(customer_data)

```
Write a code to find the dimensions of the data
```{r}
dim(customer_data)
```

Write a code to see the head  and tail of the dataset atleast 10rows
```{r }
head(customer_data)

tail(customer_data)
```
Write a code to see the data types of the attributes 
```{r }
str(customer_data)

```
Write a code to remove the unuseful variables and store the data 
```{r }
sum(is.na(customer_data))
```
Write a code to observe the variables and convert them into the required formats
```{r}
customer_data$CustomerID=as.factor(customer_data$CustomerID)
customer_data$FavoriteChannelOfTransaction=as.character(customer_data$FavoriteChannelOfTransaction)
customer_data$FavoriteGame=as.character(customer_data$FavoriteGame)
```
Check the changes again using str() command
```{r }
str(customer_data)

```
Write a Code to seperate the target
```{r}
train_rows <- sample(x = 1:nrow(customer_data), size = 0.7*nrow(customer_data))
train_data <- customer_data[train_rows, ]

test_data <- customer_data[-train_rows, ]


```
Write a code to get the numerical attributes out and store them as data_num
```{r }
data_num=customer_data[,c("City","NoOfChildren","MinAgeOfChild","MaxAgeOfChild","Tenure","FrquncyOfPurchase","NoOfUnitsPurchased","FrequencyOFPlay","NoOfGamesPlayed","NoOfGamesBought","TotalRevenueGenerated")]
str(data_num)

```
Write a code for Normalizing the data_num
```{r}
library(caret)
std_model <- preProcess(train_data[,!names(train_data) %in% c("TotalRevenueGenerated ")], method = c("center","scale"))
train_data[, !names(