---
title: "20170527_Batch29_CSE7112c_R_lab01_Basics02"
author: "Insofe"
date: "May 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir="D:\\Batch 29\\Batch29_R\\INPUTS")
```


##clearing work space

```{r cars}
rm(list=ls(all=TRUE))
```

##know your working directory
```{r}
getwd()
```

##set your working directory in which you have input files
```{r}
#setwd("C://Users/Shiva/Desktop/CPEE/Batch29/INPUTS")
```

##Reading different types of files(basic commands)
1.Reading the csv files into R environment

```{r}
Grade<-read.csv("Grade1.csv",header=T,sep=",")
dim(Grade)
```

2.Reading other formats we use read.table command

```{r}
read<-read.table("greek.txt",sep="\t",header=T)

```

3.Reading Excel files:package required is XLConnect

```{r}
#install.packages("XLConnect")
require(XLConnect) #library(XLConnect) # To loas an add-on package
wb<-loadWorkbook("Book3.xlsx")
sheet<-readWorksheet(wb,"Sheet1",header=T)
```

Explore on your own there will be more functions availbale for multiple types of files and formats



##saving you ouput files(basic commands)
1.writing into a CSV file
```{r}
write.csv(Grade,"data.csv", row.names=F)


```
2.writing into Excel
```{r}
library(xlsx)
write.xlsx(sheet,"OUTPUTSwrite.xlsx",row.names=F)

```

##Conditional statements

1.if Conditional Statement

"
if(boolean_expression) {
   // statement(s) will execute if the boolean expression is true.
}

"
```{r}
x <- 30
if(is.numeric(x)) {
  print("X is an numeric")
}
age<-45
if(age>45){
  print("middle-age person")
}
```

2.if-else Conditional Statement

if(boolean_expression) 
{
   // statement(s) will execute if the boolean expression is true.
} 
else 
{
// statement(s) will execute if the boolean expression is false.
}

```{r}
x <- c("what","is","truth")

if("Truth" %in% x) {
  print("Truth is found")
} else {
  print("Truth is not found")
}
```

##Loops
For Loop
"
for (value in vector) {
  statements
}
"

```{r}
v <- LETTERS[1:4]
v
for ( i in v) {
  print(i)
}

for(i in 1:length(v)){
  
  print(v[i])
}
```

While loop
"
while (test_expression) {
   statement
}
"
```{r}
v <- c("Hello","while loop")
cnt <- 0

while (cnt < 4) {
  print(v)
  cnt = cnt + 1
}
```


"For example I want to categorize the students into 3 bins based on the overall percentage as follows
if pct < 40 then C, if 40 < pct < 60,B and if greater than 60 then A
"
Using for loop
```{r}


Grade$class<-NA
for(i in 1:nrow(Grade)){
  if(Grade$OverallPct1[i]<40){
    Grade$class[i]<-"C"
  }else{
    if(Grade$OverallPct1[i]<60){
      Grade$class[i]<-"B"
    }else{
      Grade$class[i]<-"A"
    }
  }
}
```

Doing the same using ifelse
```{r}
Grade$Class1<-ifelse(Grade$OverallPct1<40,"C",ifelse(Grade$OverallPct1<60,"B","A"))

```

Again Giving multiple conditions in the if or ifelse statements
```{r}
Grade$class2<-ifelse(Grade$OverallPct1<40 | Grade$Math1<60, "C",
                     ifelse (Grade$OverallPct1<60 |Grade$Math1<80,"B","A")) ##for "and" use "&"
```


##Writing function in R

1.built in functions
returns sum
```{r}
v<-c(1,2,3,4,5)
sum(v)
```
returns mean and standard deviation
```{r}
v<-c(1,2,3,4,5)
mean(v)
sd(v)
```

2.user defined functions
"
function_name <- function(arg_1, arg_2, ...) {
   Function body 
}
"
```{r}
square<-function(x){
  return(x^2)
}
square(3)
square(c(3,2))
```

you can also apply a function on data frame
```{r}
y<-data.frame(A=c(1,2,3,4),B=c(5,6,7,8))
y
square(y)
```


##Some useful functions used in data manipulations
1.apply

```{r}
attach(mtcars)
data<-mtcars
str(data)
summary(data)
```

suppose if we Want to find min value for each column in mtcars dataset then
```{r}
data=mtcars
min(data$mpg)
a<-apply(data,2,min) #This generates the min values for each 

```

writing the output of previous code to a data frame
```{r}
A<-apply(data[,2:11],2,min)
A<-data.frame(min=apply(data[,2:11],2,min))
B<-apply(data[,2:11],2,max)
```


##We can find the stats for each of the variable separately
##If we want to have all the stats in a data frame we can write a customize function for this
```{r}
stat<-function(x){
  "Mean"=mean(x)
  "Min"=min(x)
  "Max"=max(x)
  A<-data.frame(Min,Mean,Max)
  return(A)
}
stats<-apply(data[,2:11],2,FUN=stat) ##Observe the ouptput of apply.. it is a list
class(stats)
result<-do.call(rbind,stats)
```

2.lapply
```{r}
l<-lapply(data[,2:11],mean) 
```
3.tapply
```{r}
tappy<-tapply(mtcars$mpg,mtcars$cyl,mean) # takes one function and gives the values and not a dataframe
tappy ##This gives out the mean mileage for each cylinder types
```

##Subsetting

Subset on vectors

```{r}
v<-c(1,2,3,4,5)
v[v>3]  #Output all elements greater than 3
```

Subset on matrices and data frames
a. Calling by cell positions
```{r}
#library(dplyr)
data<-mtcars
data1<-data[,2:11]
data1<-data[1:10,2:11]
```

b. By using column names
```{r}
data1<-data[,c("mpg","cyl")]
```

c. Using a subset function ##from help identify the argument to be given
```{r}
data1<-subset(data,mpg>25,select=mpg:carb)
```

d. The same dataframe can be obtained in another way
```{r}
data1<-data[mpg>25,]
```

##Multiple conditions can be given using "&" or "|"
```{r}
data<-mtcars
attach(data)
data2<-data[mpg>25 & hp>75,]
data2<-subset(data,mpg>25 | gear==5,select=mpg:carb)
filter(data,mpg>25 | gear==5)
```

Using which.max
```{r}
data[which.max(mpg),]
```

Using which.min
```{r}
data[which.min(mpg),]
```

Using which

```{r}
data[which(data$mpg==max(data$mpg)),]
data[which(row.names(data) %in% c("Mazda RX4","Datsun 710")),]
```

