---
title: B29 Activity-2 27th May 2017"
output: html_notebook
author: Your Name
---
```{r}
rm(list=ls(all=TRUE))
```

```{r}
install.packages("infotheo")
```

```{r}
setwd("C:/class/27052017_A1-kirankumarpatro-master/Data")
getwd()
```

##### 1) Create a function which takes vector as an input and return sine of those values

```{r}
deg2rad <- function(deg)
{
  (deg * pi/180)
}
sine <- function(x)
{
  return(sin(deg2rad(x)))
}
z <- c(0,30,45,60,90)
p <- sine(z)
```

##### 2) Take data set CustomerData1 and remove NA values and store that into variable mydata. Check structure of the data

```{r}
mydata <- read.csv("CustomerData1.csv",header = T,sep =",")
mydata <- na.omit(mydata)
```



##### 3) First column is not useful, so remove and store result back into variable mydata

```{r}
mydata <- mydata[,2:14]
```

#####4) Convert the variable City into factor
```{r}
mydata$City <- as.factor(mydata$City)
```

##### 5) For mtcars dataset, find mean of hp for each group of gear

```{r}
mtcars <- head(mtcars)
mtcars1 <- aggregate(mtcars$hp,by=list(mtcars$gear),FUN=mean,na.rm=T)
```

##### 6) Plot histogram of variable Revenue in mydata

```{r}
histogram <- hist(mydata$Revenue, col = "Yellow", xlab = "Revenue", ylab = "Frequency", main = "Revenue Plot")
```

##### 7) Create dummy variables for City which has been converted into a factor. Put the dummy variables and other variables in datafram mydata2

```{r}
library(dummies)
str(mydata)
citydummyvar <- dummy(mydata$City)

mydata2 <- data.frame(mydata,citydummyvar)
```

##### 8 ) Remove original City variable from the mydata2

```{r}
mydata2 <- mydata2[,2:15]
```

##### 9) Use discretize function from infotheo library and use it to discretize Revenue variable in 5 bins. Use equalfreq as value for the paramter disc
```{r}
library(infotheo)
Revenuebin <- discretize(mydata$Revenue, disc="equalfreq",nbins=5)
```

##### 10) Use tapply function to get the median of each of the 5 bins of Revenue

```{r}
medianrevenue <- tapply(mydata$Revenue,Revenuebin,FUN = median, na.rm=T)
```


